---
title: "Monday Data Viz: Exploring Your Own Data, Fitbit series"
author: "Jessica Hoehner"
date: "`r Sys.Date()`"
output: html_document
---

Sources: 
https://mrkaye97.github.io/fitbitr/index.html
https://hydroecology.net/getting-detailed-fitbit-data-with-r/
https://health.gov/sites/default/files/2019-10/PAG_ExecutiveSummary.pdf

```{r setup, include=FALSE}
pacman::p_load("fitbitr", "tidyverse", "ggplot2", "keyring", 
               "viridis", "lubridate", "here", "forcats", 
               "fitbitViz", "glitr", "extrafont", "ggtext", "glue")

token <- generate_token(
  client_id = key_get("client_id"),
  client_secret = key_get("client_secret"),
  callback = key_get("callback")
)

output <- here("fitviz/output")

```

## How often am I meeting the CDC's reccomended guidelines for adult physical activity?

1. 150-300 minutes per week of moderate intensity activity OR
2. 75-150 minutes per week of vigorous intensity activity OR
3. 75-150 minutes per week of moderate intensity + vigorous intensity activity 

First, I pull in the minutes I have been "fairly" and "very" active over the
last several weeks. "Fairly" and "Very" correspond with moderate and vigorous 
activity respectively. Next, I pivot the data to create a column with 
"activity level" and summarize the daily data by week.

```{r activity1}

start_date <- as_date("2022-08-02")
end_date <- as_date("2022-09-02")

fair <- minutes_fairly_active(start_date, end_date)
very <- minutes_very_active(start_date, end_date)

total_activity <- fair %>%
  left_join(very, by = "date") %>%
  mutate(
    across(.cols = (starts_with("minutes")), 
           ~ as.numeric(.)), 
    week = as_date(floor_date(date, unit = "week"))) %>%
   pivot_longer(., 
                starts_with("minutes"), 
                            names_to = "intensity", 
                            values_to = "minutes") %>%
  select(-date) %>%
  group_by(week) %>%
  mutate(sum_combo = sum(minutes)) %>%
  group_by(week, intensity) %>%
   mutate(
     intensity = stringr::str_remove(intensity, "minutes_"),
     sum = sum(minutes), 
     mean = mean(minutes),
     median = median(minutes)) %>%
  select(-minutes) %>%
  distinct()
 
```

## Part 1: Sum

```{r overall_plot, echo=FALSE}

ggplot(total_activity %>%
         mutate(
         goal = as.numeric(150))) +
  si_style() +
  geom_col(aes(x = week, y = sum_combo, fill = moody_blue), 
           position = "dodge") +
    geom_col(aes(x = week, y = goal, fill = moody_blue), alpha = 0.3,
           position = "dodge") +
  scale_y_continuous(breaks = c(0, 25, 50, 75, 100, 125, 175, 
                                150, 175, 200, 225, 250, 
                                275, 300, 325, 350, 375, 400), 
                     limits = c(0,400)) +
  scale_fill_manual(values = c(moody_blue, moody_blue)) +
  labs(
    x = NULL, 
    y = NULL,
    fill = NULL,
    legend = NULL, 
    alpha = NULL, 
    caption = "Source: Fitbit API, my data",
    # title = glue("<i style='color:{usaid_black};'>Minimum Goals Exceeded Each Week Device Was Worn,<br> Maximum Goals Exceeded Twice</i>") %>% toupper, 
    subtitle = glue("<i style='color:{usaid_darkgrey};'>I exceeded the <i style='color:{moody_blue};'>150 minute goal</i> <i style='color:{usaid_darkgrey};'>twice over the last five weeks</i>")) +
  theme(legend.position = "none", 
        text = element_text(family = "Source Sans Pro"),
        plot.title = element_markdown(), 
        plot.subtitle = element_markdown()) +
  annotate("text", 
           x = as_date("2022-08-07"), y = 275, 
           label = "1.7x max. goal", 
           color = usaid_darkgrey) +
  annotate("text", 
           x = as_date("2022-08-28"), y = 350, 
           label = "2.2x max. goal", 
           color = usaid_darkgrey)

si_save("overall_activity.png")

```
Looking at the sum of all active minutes overall, I seem like a very active person.
I met the minimum goal of 75 minutes of moderate + vigorous activity per week 
each week (except for the week I did not wear my device, Aug 22) and even exceeded 
the maximum goal of 150 minutes twice over the last five weeks.

For the first week I exceeded the maximum goal (Aug 08), I was attending a professional conference that had me commuting daily and running from room to room for most of each day that week. During the week of August 29th, I have started running more regularly
again and ran about 6.5 miles that week.

Let's look at activity by intensity to see how I am doing with regard to moderate
vs. vigorous activity.

```{r activity_plot, echo=FALSE}

ggplot(total_activity %>%
         mutate(goal = as.numeric(case_when(
                      intensity == "fairly_active" ~ 150, 
                      intensity == "very_active" ~ 75)))) +
  si_style() +
  geom_col(aes(x = week, y = sum, fill = intensity, group = intensity), 
           position = "dodge") +
  geom_col(aes(x = week, y = goal, 
               fill = intensity, group = intensity), alpha = 0.3,
           position = "dodge") +
  scale_fill_manual(values = c(golden_sand,burnt_sienna),
                    labels = NULL) +
  scale_y_continuous(breaks = c(0, 25, 50, 75, 100, 125, 175, 
                                150, 175, 200, 225, 250, 
                                275, 300, 325, 350, 375, 400), 
                     limits = c(0,400)) +
  labs(
    x = NULL, 
    y = NULL,
    fill = NULL,
    legend = NULL, 
    alpha = NULL, 
    caption = "Source: Fitbit API, my data",
    title =  glue("<i style='color:{usaid_black};'>A 60 MPH COUCH POTATO</i>"), 
    subtitle = glue("<i style='color:{usaid_darkgrey};'>
    While I have met the minimum goal for</i> <i style='color:{burnt_sienna};'>vigorous actvity</i> <i style='color:{usaid_darkgrey};'>most weeks,</i> <br> 
    <i style='color:{usaid_darkgrey};'>I have only met the minimum goal for</i> 
    <i style='color:{golden_sand};'>moderate activity</i>
    <i style='color:{usaid_darkgrey};'>once in the last five weeks</i>")) +
  theme(legend.position = "none", 
        text = element_text(family = "Source Sans Pro"),
        plot.title = element_markdown(), 
        plot.subtitle = element_markdown())

si_save("activity_by_intensity.png")

```

Over the last five weeks I have met the minimum goal for vigorous activity (75 minutes)
most weeks and even exceeded the maximum goal of 150 minutes (by one minute!) while at the conference. Week to week I appear to struggle with meeting the minimum goal of 150 minutes of moderate activity and the only met this goal the week I started running 
regularly again (Aug 29)

Looking at how inconsistently I am moderately active, this suggests to me 
that most of every day in a given week I am fairly sedentary but that there
are 1-2 days each week on which I am very active. I am curious how this will 
show up in my average minutes of weekly activity which we'll look at in Part II!