---
title: "Monday Data Viz: Exploring Your Own Data, Fitbit series"
author: "Jessica Hoehner"
date: "`r Sys.Date()`"
output: html_document
---

Sources: 
https://mrkaye97.github.io/fitbitr/index.html
https://hydroecology.net/getting-detailed-fitbit-data-with-r/
https://health.gov/sites/default/files/2019-10/PAG_ExecutiveSummary.pdf

```{r setup, include=FALSE}
pacman::p_load("fitbitr", "tidyverse", "ggplot2", "keyring", 
               "viridis", "lubridate", "here", "forcats", 
               "fitbitViz", "glitr", "extrafont", "ggtext", "glue")

token <- generate_token(
  client_id = key_get("client_id"),
  client_secret = key_get("client_secret"),
  callback = key_get("callback")
)

output <- here("fitviz/output")

```

## How often am I meeting the CDC's reccomended guidelines for adult physical activity?

1. 150-300 minutes a week of moderate-intensity activity OR
2. 75-150 minutes a week of vigorous-intensity activity OR
3. 75-150 minutes of moderate-intensity + vigorous-intensity activity 

First, I pull in the minutes I have been "fairly" and "very" active over the
last several weeks. "Fairly" and "Very" correspond with moderate and vigorous 
activity respectively. Next, I pivot the data to create a column with 
"activity level" and summarize the daily data by week.

```{r activity1}

start_date <- as_date("2022-07-17")
end_date <- as_date("2022-08-24")

fair <- minutes_fairly_active(start_date, end_date)
very <- minutes_very_active(start_date, end_date)

total_activity <- fair %>%
  left_join(very, by = "date") %>%
  mutate(
    across(.cols = (starts_with("minutes")), 
           ~ as.numeric(.)), 
    week = as_date(floor_date(date, unit = "week"))) %>%
   pivot_longer(., 
                starts_with("minutes"), 
                            names_to = "intensity", 
                            values_to = "minutes") %>%
  filter(minutes != 0) %>%
  select(-date) %>%
  group_by(week) %>%
  mutate(sum_combo = sum(minutes)) %>%
  group_by(week, intensity) %>%
   mutate(
     intensity = stringr::str_remove(intensity, "minutes_"),
     sum = sum(minutes), 
     mean = mean(minutes),
     median = median(minutes)) %>%
  select(-minutes) %>%
  distinct()
 
```

## Part 1: Sum

```{r overall_plot, echo=FALSE}

ggplot(total_activity %>%
         mutate(
         goal = as.numeric(75))) +
  si_style() +
  geom_col(aes(x = week, y = sum_combo, fill = moody_blue), 
           position = "dodge") +
    geom_col(aes(x = week, y = goal, fill = moody_blue, alpha = 0.8), 
           position = "dodge") +
  scale_y_continuous(breaks = c(0, 25, 50, 75, 100, 125, 175, 
                                150, 175, 200, 225, 250, 
                                275, 300, 325, 350, 375, 400), 
                     limits = c(0,400)) +
  scale_fill_manual(values = c(moody_blue, moody_blue)) +
  labs(
    x = NULL, 
    y = "Minutes",
    fill = NULL,
    legend = NULL, 
    alpha = NULL, 
    caption = "J. Hoehner",
    title = glue("<i style='color:{usaid_black};'>MINIMUM RECOMMENDATIONS MET 
                 EACH WEEK, EXCEEDED SOME WEEKS</i>"), 
    subtitle = glue("<i style='color:{usaid_darkgrey};'>Over the last five weeks 
    there have been two weeks in which I have</i> <br> 
    <i style='color:{usaid_darkgrey};'>exceeded the maximum recommended number 
    of active minutes per week by 1.7-2x</i>")) +
  theme(legend.position = "none", 
        text = element_text(family = "Source Sans Pro"),
        plot.title = element_markdown(), 
        plot.subtitle = element_markdown()) +
  annotate("text", 
           x = as_date("2022-08-07"), y = 275, 
           label = "1.7x max. goal") +
  annotate("text", 
           x = as_date("2022-07-17"), y = 325, 
           label = "2x max. goal")

si_save("overall_activity.png")

```
Looking at the sum of all active minutes overall, I seem like a very active person.
I have been meeting the minimum goal of 75 minutes of moderate + vigorous activity 
per week and even exceeded the maximum goal of 150 minutes twice over the last five weeks.

I imagine I was especially active the first week documented because
I was excited to start wearing my Fitbit again and for the other week I exceeded
the goal (Aug 08), I was attending a professional conference that had me 
commuting daily and running from room to room for most of each day that week. 

Let's look at activity by intensity to see how I am doing with regard to moderate
vs. vigorous activity.

```{r activity_plot, echo=FALSE}

ggplot(total_activity %>%
         mutate(goal = as.numeric(case_when(
                      intensity == "fairly_active" ~ 150, 
                      intensity == "very_active" ~ 75)))) +
  si_style() +
  geom_col(aes(x = week, y = sum, fill = intensity, group = intensity), 
           position = "dodge") +
  geom_col(aes(x = week, y = goal, alpha = 0.05, 
               fill = intensity, group = intensity), 
           position = "dodge") +
  scale_fill_manual(values = c(golden_sand,burnt_sienna),
                    labels = NULL) +
  scale_y_continuous(breaks = c(0, 25, 50, 75, 100, 125, 175, 
                                150, 175, 200, 225, 250, 
                                275, 300, 325, 350, 375, 400), 
                     limits = c(0,400)) +
  labs(
    x = NULL, 
    y = "Minutes",
    fill = NULL,
    legend = NULL, 
    alpha = NULL, 
    caption = "J. Hoehner",
    title =  glue("<i style='color:{usaid_black};'>A 60 MPH COUCH POTATO</i>"), 
    subtitle = glue("<i style='color:{usaid_darkgrey};'>
    While I have met the minimum goal for</i> <i style='color:{burnt_sienna};'>vigorous actvity</i> <i style='color:{usaid_darkgrey};'>most weeks,</i> <br> 
    <i style='color:{usaid_darkgrey};'>I have only met the minimum goal for</i> 
    <i style='color:{golden_sand};'>moderate activity</i>
    <i style='color:{usaid_darkgrey};'>once in the previous five weeks</i>")) +
  theme(legend.position = "none", 
        text = element_text(family = "Source Sans Pro"),
        plot.title = element_markdown(), 
        plot.subtitle = element_markdown())

si_save("activity_by_intentsity.png")

```

Over the previous five weeks I have met the minimum goal for vigorous activity (75 minutes) 
most weeks and even exceeded the maximum goal of 150 minutes (by one minute!) while at the conference. 
Week to week I appear to struggle with meeting the minimum goal of 150 minutes of moderate 
activity and the only met this goal the first week I wore the Fitbit.

Looking at how inconsistently I am moderately active, this suggests to me 
that most of every day in a given week I am fairly sedentary but that there
are 1-2 days each week on which I am very active. I am curious how this will 
show up in my average minutes of weekly activity which we'll look at in Part II!